---
title: "hw04-MaunishUBC"
output: github_document
always_allow_html: yes
---

### Homework 04: Tidy data and joins ###

Install packages for analysis
```{r warning=FALSE}
library(gapminder)
head(gapminder)
library(tidyverse)
library(kableExtra)
```


### Data Reshaping: Activity#2 ### 


#### Break the dataset so that it has information for 3 countries: Canada, India and Rwanda ####

```{r}
life_exp_countries <- gapminder %>% 
  filter(country %in% c("Canada", "India", "Rwanda")) %>% 
  select(country, year, lifeExp)
life_exp_countries
```


#### Let's reshape this such that each row is a year and each column is a country ####

```{r}
life_exp_countries_untidy <- spread(life_exp_countries, key = "country", value = lifeExp)
life_exp_countries_untidy %>% kable(row.names = TRUE, align = 'c') %>% 
  kable_styling()
```


#### Generate scatter plots of one country against another ####

```{r}
life_exp_countries_untidy %>% ggplot(aes(India, Canada)) +
  geom_point() + 
  geom_path(arrow = arrow())
```

```{r}
life_exp_countries_untidy %>% ggplot(aes(Rwanda, Canada)) +
  geom_point() + 
  geom_path(arrow = arrow())
```
```{r}
life_exp_countries_untidy %>% ggplot(aes(Rwanda, India)) +
  geom_point() + 
  geom_path(arrow = arrow())
```


### Join Prompts: Activity 1 ###

```{r}
head(gapminder) 
```

```{r warning=FALSE}
capital_info <- read_csv("http://techslides.com/demos/country-capitals.csv")
capital_info
```

To make the capital_info dataset comparable to gapminder rename *CountryName* as **country** and *CapitalName* as **capital**

```{r}
capital_info_tidy <- capital_info %>% rename(country = CountryName, capital = CapitalName, capital_latitude = CapitalLatitude) %>% 
  select(country, capital, capital_latitude)

capital_info_tidy %>% kable(row.names = TRUE) %>% 
  kable_styling()

```


Based on the capital's latitude a country will be assigned to the north or south hemisphere. If hemisphere = 1  then it is North Hemisphere otherwise country is in south hemisphere

```{r}

capital_info_hemisphere <- capital_info_tidy %>% 
  mutate(hemisphere = if_else(capital_latitude > 0,1,0))
capital_info_hemisphere %>% kable(row.names = TRUE) %>% 
  kable_styling()
```

#### Let's combine the gapminder and the capital_info_hemisphere datasets using *inner join* first
```{r}
gap_minder_with_capital_inner_join <- inner_join(gapminder, capital_info_hemisphere, by = "country")
gap_minder_with_capital_inner_join
```

Note that this matrix has 1584 rows whereas gapminder has 1694 rows, I suspect that's due to difference in the country names in the two datasets

```{r}
unique(gap_minder_with_capital_inner_join$country)
unique(gapminder$country)
unique(capital_info_hemisphere$country)
```

So, these datasets are indeed quite different with respect to the country info they carry. The capital_info dataset seems to be more exhaustive than the gapminder dataset. 

Let's look at the differnces better using *anti_join*
```{r}
anti_join(capital_info_hemisphere, gapminder, by = "country")
```

As gapminder is our primary focus let's see how does it look with the added information from the capital_info dataset, for this we will use *left_join*

```{r}
left_join(gapminder, capital_info_hemisphere, by = "country")
```

For all the data combined into a single dataset we will use *full join* 
```{r}
gapminder_metadata <- full_join(gapminder, capital_info_hemisphere, by = "country")
gapminder_metadata
```

So, what countries are unique to the capital_info dataset?
```{r}
subset(gapminder_metadata, is.na(year)) %>% 
  kable(row.names = TRUE) %>% 
  kable_styling()
```

What countries are unique to the gapminder dataset? 
```{r}
unique(subset(gapminder_metadata, is.na(capital))$country)
```

